Formalizing Commutative Algebra in Coq: \\ Nakayama's Lemma\thanks{Source code for this work is available on the following site: \url{https://github.com/ku-sldg/algebra}}

### 2023-02-02 nakayama/paper.tex
Edited the section on the formal proof, minor edits throughout exposition on Coq, and added Conclusion

\author{ 
	Andrew Cousino \\ {\ttfamily acousino@ku.edu}
	\and 
	Emily E.\ Witt\thanks{Witt acknowledges support from NSF CAREER Award
    DMS-1945611, the Keeler Intra-University Professorship from the University of Kansas, and the 2022-23 Ruth I.\, Michler Memorial Prize from the
    Association for Women in Mathematics.} \\ {\ttfamily witt@ku.edu}
  \and 
	Perry Alexander \\ {\ttfamily palexand@ku.edu}
}

@@

Our goal is to newly formalize theoretical, rather than computational,
commutative algebra in Coq. We formally prove \emph{Nakayama's Lemma}
\cite{nakayama-1951, azumaya}, an essential result in the field. In doing so,
we formalize algebraic structures that are fundamental to higher-level algebra,
such as \emph{local rings} and \emph{modules over commutative rings}, and
\emph{quotient rings and modules}. Rather than build upon some of the basic
objects from abstract algebra, such as groups and rings, that have been
formalized in Coq, e.g., in the \emph{Mathematical Components Library}
\cite{mathcomp}, we start from scratch.
\todo{To do: Verify whether MathComp only formalized finite algebraic
  structures. Drew is about as certain as he can be that this is the case.}
The theory, including the formalization of all algebraic structures, makes up
approximately $100$ kB and 3300 lines of code.

@@

\begin{figure}[t]
  \caption{The hierarchy of our algebraic structures}
  \includegraphics[width=.4\textwidth]{algebraic-structures.pdf}
  \centering
\end{figure}

\subsection{Semigroups} Our foundation begins by defining a semigroup class, which declares a binary
operation to be associative. From here, we build up through monoids, which
introduce identities, to groups, which introduce inverses. Note the double
equals ``$==$'' appearing in these definitions is notation for an arbitrary
equivalence relation over the group's carrier set, which acts as equality.

\begin{minted}{coq}
  Infix "==" := equiv (at level 60, no associativity).
  Class Semigroup := {
    semigroup_assoc:
      forall (a b c: Carrier),
        a <o> b <o> c == a <o> (b <o> c);
  }.
  Class Monoid := {
    monoid_semigroup :> Semigroup equiv op;
    monoid_ident_l:
      forall (a: Carrier), ident <o> a == a;
    monoid_ident_r:
      forall (a: Carrier), a <o> ident == a;
  }.
  Class Group := {
    group_monoid :> Monoid equiv op ident;
    group_inv_l:
      forall (a: Carrier), inv a <o> a == ident;
    group_inv_r:
      forall (a: Carrier), a <o> inv a == ident;
  }.
\end{minted}
The lines such as  ``\verb|monoid_semigroup :> Semigroup equiv op;|'' simply coerce
the monoid typeclass into a semigroup.

\subsection{Algebraic quotients}
While in the end, our formal proof does not call upon quotients of
algebraic structures, alternative proofs of Nakamaya's Lemma take advantage of 
 the structures quotient rings and quotient modules.  As algebraic 
 quotient structures are also fundamental to commutative algebra,
it is work pointing out that we have formalized quotients of groups, rings, and modules 
  using typeclasses, which appear to work rather smoothly.

@@

The utility--and, as a consequence, the ubiquity--of quotients in commutative algebra 
motivates our choice to use equivalence relations to define the components of a group structure. If one
were to instead use the traditional Leibniz equality, it would be difficult to
identify a quotient group with another group. However, by  defining a group in
terms of an arbitrary equivalence relation, in our theory a quotient group is 
simply defined as a group, but under an equivalence relation that is not the
usual equality. Not much is lost in adopting this convention thanks to Coq's 
rewrite tactics for setoids--types equipped with an equivalence relation---can 
still be called
upon.

\subsection{Rings and their ideals}
Moving onward, rings form the next step in our algebraic hierarchy; a ring has
two binary operations: addition, which must be commutative, and multiplication,
which need not be commutative in general. In our formulation, rings must have a
multiplicative identity. Next, we formalized the definition of a commutative
ring, further requiring commutativity of multiplication. At this point, we
formalized the notion of an ideal of a commutative ring, a subgroup of the ring
under addition that satisfies the absorption property under multiplication,
i.e., \(r a\) sits in the ideal whenever \(a\) of the ideal, regardless of the choice of 
element \(r\) of the commutative ring. We also used this to formalize the
notion of a quotient ring $R/I$, where $I$ is an ideal of a commutative ring
$R$.

Next, we formalized the definition of a  prime ideal, and then 
moved on to do the same for the notion of a maximal ideal, a proper ideal 
that is maximal with respect to inclusion. Below is the definition in Coq,
which uses \texttt{P} as the predicate for the ideal.

@@

\noindent Next, we were able to employ this definition of a maximal ideal
to formally define a local
ring, i.e., a commutative ring with a single maximal ideal.

@@

\subsection{Modules over rings}
At this point, we formalized the definition of a module over a commutative ring,
the commutative-algebraic generalization of the notion of a vector space over a
field. Nakayama's Lemma is a statement about finitely generated modules, and
hence we formalized the notion of a scalar combination of a finite
collection of elements, $u_1, \ldots, u_n$ a module $M$ over a commutative ring $R$, i.e.,
expressions of the form $r_1 u_1 + r_1 u_2 + \cdots + r_n u_n$, where each
$r_i$ is an element of $R$.

In our formalization of scalar combinations, we use  ``list'' to mean
length-parameterized lists; since we don't use the simpler kind of lists, there
are no name collisions. In our code, \texttt{M} is the type of module elements,
\texttt{R} is the type of ring elements, act as coefficients, and
\texttt{t A n} is a list whose elements are of type \texttt{A} and whose length
is \texttt{n}.

@@

Finally, we defined the submodule $IM$ of a module $M$ over a commutative ring $R$, 
where $I$ is an arbitrary ideal of $R$.
Recall that when $R$ is local and $I$ is its unique maximal ideal, this submodule 
appears in the statement of Nakayama's Lemma. 
In general, $IM$ is the set consisting of all scalar
combinations of elements of \(M\) whose coefficients are in  \(I\). We
represent this in Coq as a predicate over \(M\).

@@

\section{Constructing the Formal Proof}
Not only to aid in the formalization of Nakayama's lemma, but also to

Beyond formalizing the relevant structures from higher algebra, we also
formally establish some basic theory from the field in Coq. For instance, 
consider the notion of a \emph{unit} of a
commutative ring $R$, which is an element $a \in R$ with a multiplicative inverse,
i.e., there exists an element $a^{-1} \in R$ for which $x^{-1} \cdot x$ is the
multiplicative identity $1 \in R$.  In fact, it $x$ is a unit, its inverse is unique.

\emily[inline]{Emily: Just changed statement to \emph{if and only if} below; we should check whether we should update its formal proof}
Suppose that $I$ is an ideal of a commutative rings $R$.  
We formally establish the following: If $I$ contains a
unit $a$ if and only if $I$ is the trivial ideal, i.e., $I=R$. 
The informal proof logic is as follows:  Due to the absorption property of $I$,  
$a^{-1}  \cdot a= 1 \in I$.  Hence for every element $r$ of $R$,
$r = r \cdot 1$ is also in $I$, i.e., $I=R$. 
On the other hand, if $I = R$, then the multiplicative identity $1$ is a unit in $I$.

\subsection{Dealing with the Axiom of Choice}
Every non-unit element of a commutative ring is contained in some maximal
ideal, a fact that relies on the Axiom of Choice. 
The following standard informal
proof calls upon Zorn's lemma, which is equivalent to the Axiom of Choice, and which 
says that if every chain of elements from a partially ordered set  has an upper
bound in the set, then the set must contain at least one maximal element.

\begin{quote}
	Fix a non-unit $x$ in a commutative ring $R$, and  \(I_{1}\) to be the principal 
	ideal generated by $x$, i.e., the smallest ideal containing the element $x$,
    which consists of all $R$-multiples of it.  No element of $I_1$ is a unit, since if  
    $y \in R$ were the multiplicative inverse of some $R$-multiple
    $r \cdot x$, then $1 = y \cdot (r \cdot x) = (y\cdot r) \cdot x$, showing that $x^{-1} = y \cdot r$, 
    which contradicts the assumption that $x$ is not a unit. Hence by the statement described earlier in this section, 
    $I_1$ is proper.

   Either $I_1$ is a maximal ideal of $R$, or it is strictly contained in a maximal ideal, or there exists 
   an inifinite chain of strictly increasing
   proper ideals of the form 
    \begin{equation} \label{infinite-chain: e}
    I_{1}\subsetneq I_{2}\subsetneq I_{3}\subsetneq\cdots\subsetneq I_n \subsetneq \cdots. 
    \end{equation}  
   Indeed, suppose that $I_1$ is not maximal, and   
   assume that for some integer $n \geq 1$,  there exist proper, non-maximal ideals 
   $I_k$ of $R$, $1 < k \leq n$, such that for all such $k$,  $I_{k-1}$ is strictly contained in $I_k$.  
   Since $I_n$ is not maximal, there exists a proper ideal $I_{n+1}$ of $R$ such that $I_n \subsetneq I_{n+1}$.
      Then either $I_{n+1}$ is maximal, or we have extended the chain by one additional link.

%   
%    If \(I_{1}\) is not a maximal ideal of $R$, then by definition there exists a strictly larger
%    proper ideal \(I_{2}\) of $R$, i.e.,
%    $x \in I_{1}\subsetneq I_{2} \subsetneq R$. Likewise, if  \(I_{2}\) is again not
%    maximal, then there exists a strictly larger proper ideal
%    $I_{3}$, which, if not maximal, is strictly contained in another ideal, $I_4$, etc.
%

    It is straightforward to verify that an increasing union of ideals 
    is, by definition, an ideal containing every ideal in the union.  
Hence if, in this process, no maximal ideal is identified; i.e., we have ensured the existence of a chain of the form \eqref{infinite-chain: e},  then the increasing union of the $I_k$ is again an ideal, which contains $x$ since  $x \in I_1$.

%  Continuing in this way, if  this process terminates, we have successfully identified a
%    maximal ideal of $R$ that contains $x$, and if not, one obtains an infinite, strictly increasing chain of proper ideals of $R$, each containing $x$, 
%    \[I_{1}\subsetneq I_{2}\subsetneq I_{3}\subsetneq\cdots\subsetneq I_n \subsetneq \cdots. \]
%    It is straightforward to verify that an increasing union of ideals 
%    is, by definition, an ideal containing every ideal in the union.  Moreover, if each ideal in the union
%    contains an element $x$, then the union certainly 
%       contains \(x\) as well.

      Hence the set $X$ of all proper ideals of $R$ containing $x$, partially ordered by inclusion, 
       satisfies the property that every chain in $X$ has an upper bound in $X$.
    Therefore, Zorn's lemma ensures the existence of a maximal element of $X$, i.e., a maximal ideal of \(R\) containing \(x\).
\end{quote}

The above argument has potentially infinitely many steps, posing a Coq implementation
 problem. Indeed, suppose one were to write a proof tactic called \mintinline{coq}{generate_larger_ideal} 
allowing one to move from a non-maximal, proper ideal \(I\) that contains a fixed non-unit 
$x$ to a strictly larger ideal 
\(J \), which is again proper, and also contains \(x\).

@@

After we have overwhelmed any system with
these infinite number of calls to a single proof tactic, we still have to
invoke more tactics to finish the proof with an application of Zorn's lemma,
which finally guarantees us a maximal ideal containing \(x\). To summarize, we
are forced to run the following proof script which will never get past calling
\mintinline{coq}{generate_larger_ideal} infinitely many times.
\begin{minted}{coq}
  repeat generate_larger_ideal. (* must be called infinitely often *)
  zorns_lemma. (* called after the above non-terminating call finishes *)
\end{minted}
We chose to avoid this infinite proof script issue by including an axiom that
any non-unit \(x\) of a ring must be contained in some maximal ideal.

@@

We also used classical logic to prove that \(1 - x\) is a unit whenever \(x\)
is an element of a local ring that is not a unit. The proof is completed by way
of contradiction, and uses the rule that \(\neg\neg P\rightarrow P\).

\subsection{The Formal Statement}
We now present our formal statement of Nakayama's Lemma
in Coq, and describe a lemma called upon in our proof.

@@

To use as a building block in the formal proof of Nakayama's Lemma, we first
formally stated and proved a lemma that gives a concrete description of the
elements of the submodule $\mathfrak{m} M$  appearing in its statement. 
This lemma applies more generally, to an arbitrary finitely generated module
$M$ over a (not necessarily local) commutative ring, and any submodule of the
form $I M$, for $I$ an arbitrary ideal.

@@

This lemma's proof follows from a straightforward argument based on definitions, by
induction on the number of elements in a fixed generating set for $M$, which we
informally describe: By definition, $x \in I M$ can  be written, for some
positive integer $k$ and elements $r_i \in R$ and $w_i \in M$,
$1 \leq i \leq k$, as $x = r_1 \cdot w_1 + r_2 \cdot w_2 + \cdots +
r_k \cdot w_k$. What's more, by definition of a finite generating set $u_{1},
\dots, u_{n}$ for $M$, each $w_i$ equals $c_{i1} \cdot u_1 + c_{i2} \cdot u_2 +
\cdots + c_{in} \cdot u_n$ for appropriate choices of $c_{ij} \in R$. Hence,
inductively applying associativity, 
\begin{align*}
  x = \sum_{i=1}^{k} r_i \cdot w_i
    = \sum_{i=1}^{k} r_i \cdot \left(\sum_{j=1}^n c_{ij} \cdot  u_j\right)
    = \sum_{j=1}^{n}\sum_{i=1}^{k} r_i \cdot (c_{ij} \cdot u_j) 
    = \sum_{j=1}^{n}\left(\sum_{i=1}^{k} c_{ij} \cdot r_i\right) \cdot u_j.   
\end{align*}
By absorption in ideals, each  $c_{ij}  \cdot r_i$ is in $I$, and since ideals
are closed under addition, we inductively conclude that the coefficient $a_j
:= c_{1j} \cdot  r_1 +  c_{2j} \cdot  r_2 +  \cdots c_{kj}\cdot r_k$ of $u_j$
is also in $I$.

\subsection{The Formal Proof}
We now outline the formal proof of Nakayama's lemma, and  proceed
by induction on the number of elements in a fixed generating set for the finitely generated $R$-module
$M$. The base case in this situation is when \(M\) requires no generators, so $M$
consists solely of the empty scalar combination, i.e., the empty sum, which, by
convention, is the zero element. In other words, $M = 0$ by assumption, and
this is also precisely the conclusion of Nakayama's Lemma.  Hence the statement holds
trivially, without using the hypothesis that $M = \mathfrak{m} M$.

@@

Now, fix an $R$-module $M$ generated by $u_1, \ldots, u_{n+1} \in M$. By
assumption, \(M = \mathfrak{m} M\), and in particular, the generator $u_1$ is
an element of the submodule $\mathfrak{m} M$. The lemma described above
guarantees the existence of ring elements $a_1, \ldots, a_n \in \mathfrak{m}$
for which 
\[u_{1} = a_{1} \cdot u_{1} + a_2 \cdot u_2 + \cdots + a_{n+1} \cdot u_{n+1}.\]

Collecting the \(u_{1}\) terms on the left-hand side of the equation, applying
the existence of additive inverses in $R$, and distributivity of scalar
multiplication for modules, we see that
\begin{equation}\label{equation1: e}
(1 - a_{1}) \cdot u_{1} =
  a_{2} \cdot u_{2} + \cdots + a_{n+1} \cdot u_{n+1}\text{.}
\end{equation}
In Coq, we worked backwards to establish this fact. The comments following
each tactic offers a brief explanation of the effects on the goal.
\begin{minted}[linenos,firstnumber=157,stepnumber=5]{coq}
  assert ((Rone [+] Rminus a1) <.> u1 =M= (* (1 - a1) u1 = ... *)
    linear_combin Madd Mzero action coeffs' generatingSet').
  { setoid_rewrite (module_distrib_Radd Radd Rmul Rone
      Mequiv Madd Mzero Mminus action). (* 1 u1 + (-a1) u1 = ... *)
    setoid_rewrite (module_Rone Radd Rmul Rone
      Mequiv Madd Mzero Mminus action). (* u1 + (-a1) u1 = ... *)
    setoid_rewrite (module_minus_l Requiv Radd Rzero Rminus Rmul Rone
      Mequiv Madd Mzero Mminus action). (* u1 - a1 u1 = ... *)
    symmetry. (* ... = u1 - a1 u1 *)
    apply (group_move_r Mequiv Madd Mzero Mminus). (* ... + a1 u1 = u1 *)
    symmetry. (* u1 = ... + a1 u1 *)
    assumption. }
\end{minted}

>>>>

\author{ 
	Andrew Cousino \\ {\ttfamily acousino@ku.edu}
	\and 
	Emily E.\ Witt\thanks{Witt acknowledges support from NSF CAREER Award
    DMS-1945611, the Keeler Intra-University Professorship from the University of Kansas, and the Ruth I.\, Michler Memorial Prize from the
    Association for Women in Mathematics.} \\ {\ttfamily witt@ku.edu}
  \and 
	Perry Alexander \\ {\ttfamily palexand@ku.edu}
}

@@

Our goal is to newly formalize theoretical, rather than computational,
commutative algebra in Coq. We formally prove \emph{Nakayama's Lemma}
\cite{nakayama-1951, azumaya}, an essential result in the field. In doing so,
we formalize algebraic structures that are fundamental to higher-level algebra,
such as \emph{local rings} and \emph{modules over commutative rings}, and
\emph{quotient rings and modules}. Rather than build upon some of the basic
objects from abstract algebra, such as groups and rings, that have been
formalized in Coq, e.g., in the \emph{Mathematical Components Library}
\cite{mathcomp}, we start from scratch.
\todo{To do: Verify whether MathComp only formalized finite algebraic
  structures. Drew is about as certain as he can be that this is the case.}
The theory, including the formalization of all algebraic structures, makes up
approximately $100$ kB, and 3300 lines of code.

@@

\begin{figure}[t]
  \caption{The hierarchy of our algebraic structures in Coq}
  \includegraphics[width=.4\textwidth]{algebraic-structures.pdf}
  \centering
   \label{alg-heirarchy: F}
\end{figure}

\subsection{Semigroups} 
Appearing at the top of \Cref{alg-heirarchy: F}, our
 foundation begins by defining a semigroup class, which declares a binary
operation to be associative. From here, we build up through monoids, which
introduce identities, to groups, which introduce inverses. Note the double
equals, ``$==$'', appearing in these definitions is notation for an arbitrary
equivalence relation over the group's carrier set, which acts as equality.

\begin{minted}{coq}
  Infix "==" := equiv (at level 60, no associativity).
  Class Semigroup := {
    semigroup_assoc:
      forall (a b c: Carrier),
        a <o> b <o> c == a <o> (b <o> c);
  }.
  Class Monoid := {
    monoid_semigroup :> Semigroup equiv op;
    monoid_ident_l:
      forall (a: Carrier), ident <o> a == a;
    monoid_ident_r:
      forall (a: Carrier), a <o> ident == a;
  }.
  Class Group := {
    group_monoid :> Monoid equiv op ident;
    group_inv_l:
      forall (a: Carrier), inv a <o> a == ident;
    group_inv_r:
      forall (a: Carrier), a <o> inv a == ident;
  }.
\end{minted}
The line  ``\verb|monoid_semigroup :> Semigroup equiv op;|'' simply coerces
the monoid typeclass into a semigroup, and similar lines perform analogous functions.

\subsection{Algebraic Quotients}
While in the end, our formal proof does not call upon quotients of
algebraic structures, alternative proofs of Nakamaya's Lemma take advantage of 
 the structures quotient rings and quotient modules.  As algebraic 
 quotient structures are also fundamental to commutative algebra,
it is worth pointing out that we have formalized quotients of groups, rings, and modules 
  using typeclasses, which appear to work rather smoothly.

@@

The utility--and, as a consequence, the ubiquity--of quotients in algebra 
motivates our choice to use equivalence relations to define the components of a group structure. If one
were to instead use the traditional Leibniz equality, it would be difficult to
identify a quotient group with another group. However, by  defining a group in
terms of an arbitrary equivalence relation, in our theory a quotient group is 
simply defined as a group, but under an equivalence relation that is not the
usual equality. Not much is lost in adopting this convention, thanks to Coq's 
rewrite tactics for setoids, which are types equipped with an equivalence relation.

\subsection{Rings and their Ideals}
Moving onward, rings form the next step in our algebraic hierarchy; a ring has
two binary operations: addition, which must be commutative, and multiplication,
which need not be commutative in general. In our formulation, rings must have a
multiplicative identity. Next, we formalize the definition of a commutative
ring, further requiring commutativity of multiplication.

At this point, we formalize several algebraic structures defined in terms of commutative rings.
First is the definition of an ideal of a commutative ring, a subgroup of the ring
under addition that satisfies the absorption property under multiplication, and with this, also the 
notion of a quotient ring $R/I$, where $I$ is an ideal of a commutative ring
$R$.
We also formalize the definition of a  prime ideal, and that of a maximal ideal--a proper ideal that is maximal with respect to inclusion.
 Below is the definition of latter in Coq,
which uses \texttt{P} as the predicate for the ideal.

@@

 Next, we employ the above definition to formally define a local
ring, i.e., a commutative ring with a single maximal ideal.

@@

\subsection{Modules over Rings}
With the fundamental definitions for commutative rings and their ideals formalized, we move on 
to build our formal module theory.  
We start by formalizing the definition of a module over a commutative ring,
the commutative-algebraic generalization of the notion of a vector space over a
field. Nakayama's Lemma is a statement about finitely generated modules, and
hence we formalized the notion of a scalar combination of a finite
collection of elements, $u_1, \ldots, u_n$ a module $M$ over a commutative ring $R$, i.e.,
expressions of the form $r_1 u_1 + r_1 u_2 + \cdots + r_n u_n$, where each
$r_i$ is an element of $R$.

In our formalization of scalar combinations, we use  ``list'' to mean
length-parameterized lists; since we don't use the simpler kind of lists, there
are no name collisions. In the code excerpted below, \texttt{M} is the type of module elements,
\texttt{R} is the type of ring elements, acting as coefficients, and
\texttt{t A n} is a list whose elements are of type \texttt{A} and whose length
is \texttt{n}.

@@

Finally, we define the submodule $IM$ of a module $M$ over a commutative ring $R$, 
where $I$ is an arbitrary ideal of $R$.
Recall that when $R$ is local and $I$ is its unique maximal ideal, this submodule 
appears in the statement of Nakayama's Lemma. 
In general, $IM$ is the set consisting of all scalar
combinations of elements of \(M\) whose coefficients are in  \(I\). We
represent this in Coq as a predicate over \(M\).

@@

\section{Constructing the Formal Proof} \label{formal-proof: S}
Not only to aid in the formal proof of Nakayama's lemma, but also to add to the body 
of formalized theory in commutative algebra, along with our formal definitions of algebraic structures,
we also  
 establish some basic theory of these objects in Coq.
 For instance, 
consider the notion of a \emph{unit} of a
commutative ring $R$, which is an element $a \in R$ with a multiplicative inverse,
i.e., there exists an element $a^{-1} \in R$ for which $x^{-1} \cdot x$ is the
multiplicative identity $1 \in R$.  In fact, it $x$ is a unit, its inverse is unique.

Suppose that $I$ is an ideal of a commutative ring $R$.   Then if $I$ is the trivial ideal, i.e., $I = R$, then it, of course, contains a unit, namely $1$. 
We formally establish the converse of this statement: If $I$ contains a
unit $a$, then $I=R$. 
The informal proof logic is as follows:  Due to the absorption property of $I$,  
$a^{-1}  \cdot a= 1 \in I$.  Hence for every element $r$ of $R$,
$r = r \cdot 1$ is also in $I$, i.e., $I=R$. 
On the other hand, if $I = R$, then the multiplicative identity $1$ is a unit in $I$.

We also use classical logic to prove that \(1 - x\) is a unit whenever \(x\)
is an element of a local ring that is not a unit. The proof is completed by way
of contradiction, and uses the rule that \(\neg\neg P\rightarrow P\).

\subsection{Dealing with the Axiom of Choice}

\emily[inline]{Emily says:  There was an issue with our argument below, so I have updated it, and need to point to 
quotients to deal with the fact that we want a maximal ideal \emph{that contains a fixed element}.
}

Every non-unit element of a commutative ring is contained in some maximal
ideal, a fact that relies on the Axiom of Choice. 
This fact can be derived from from the weaker statement that every commutative ring contains a maximal ideal:
Indeed, assume this, and fix a non-unit $x$ of a commutative ring $R$.  Then the quotient $R/\langle x \rangle$ of $R$ modulo 
the principal ideal generated by $x$ is a commutative ring, and so contains a maximal ideal, which corresponds precisely to a 
maximal ideal of $R$ containing $x$ by   
the lattice isomorphism theorem for rings.

The following standard informal proof of the fact that every commutative rings contains a maximal ideal calls upon Zorn's lemma, 
which is equivalent to the Axiom of Choice (assuming Zermelo-Fraenkel Set Theory), and which 
says that if every chain of elements from a partially ordered set  has an upper
bound in the set, then the set must contain at least one maximal element.

\begin{quote}
Fix a commutative ring $R$.  Since $1 \neq 0$ in $R$, the zero ideal is proper.  
If the zero ideal is maximal, then we have identified a maximal ideal of $R$. 
If not, then there must exist a proper ideal 
$I_1$ of $R$ containing more than just the element $0$.

If $I_1$ is maximal, then we have found a maximal ideal, but if not, then there exists a proper ideal 
$I_2$ of $R$ that strictly contains $I_1$.  
Likewise, if $I_2$ is maximal, we are done, and if not, there exists a proper ideal $I_3$ such that $I_3 \supsetneq I_2$.
Continuing this pattern, we either identify a maximal ideal of $R$, or we construct an infinite chain of proper ideals of $R$:
    \begin{equation} \label{infinite-chain: e}
    I_{1}\subsetneq I_{2}\subsetneq I_{3}\subsetneq\cdots\subsetneq I_n \subsetneq \cdots. 
    \end{equation}

       More precisely, assume by induction that  there exist proper, non-maximal ideals 
   $I_k$ of $R$, $1 < k \leq n$, such that each $I_{k-1}$ is strictly contained in $I_k$.  
   Since $I_n$ is not maximal, there exists a proper ideal $I_{n+1}$ of $R$ such that $I_n \subsetneq I_{n+1}$.
      Then either $I_{n+1}$ is maximal, or we have extended the chain by one additional link.

It is straightforward to verify, by definition, 
         that an increasing union of ideals 
    is again an ideal.
Hence if, in this process, no maximal ideal is identified, i.e., we have ensured the existence of a chain of the form \eqref{infinite-chain: e},  then the increasing union of the $I_k$ is again an ideal.

Let $X$ denote the set of all proper ideals of $R$, partially ordered by inclusion.
We have shown that every chain in $X$ has an upper bound in $X$, so  Zorn's lemma ensures the existence of a maximal element of $X$, i.e., a maximal ideal of \(R\). 
\end{quote}

This argument potentially requires infinitely many steps, posing an implementation
 problem. Indeed, suppose one were to write a Coq proof tactic called \mintinline{coq}{generate_larger_ideal}
 that allows one to move from a non-maximal, proper ideal \(I\) that contains a fixed non-unit 
$x$ to a strictly larger ideal 
\(J \), which is again proper, and also contains \(x\).

@@

Even if it were possible not to overwhelm a system with
these infinitely many calls to a single proof tactic, one would still have to
invoke more tactics to finish the proof with an application of Zorn's lemma,
which finally guarantees us a maximal ideal containing \(x\). To summarize, we
are forced to run the following proof script which will never get past calling
\mintinline{coq}{generate_larger_ideal} infinitely many times.
\begin{minted}{coq}
  repeat generate_larger_ideal. (* must be called infinitely often *)
  zorns_lemma. (* called after the above non-terminating call finishes *)
\end{minted}

We choose to avoid this infinite proof script issue by including an axiom that
any non-unit \(x\) of a ring must be contained in some maximal ideal.

@@

\subsection{The Formal Statement} \label{formal-statement: SS}
We now present our formal statement of Nakayama's Lemma
in Coq, and describe a lemma called upon in our proof.

@@

To use as a building block in the formal proof of Nakayama's Lemma, we 
formally state and prove a lemma that gives a concrete description of the
elements of the submodule $\mathfrak{m} M$  appearing in its statement. 
This lemma applies more generally, to an arbitrary finitely generated module
$M$ over a (not necessarily local) commutative ring, and any submodule of the
form $I M$, for $I$ an arbitrary ideal.

@@

This lemma's proof follows from a straightforward argument based on definitions, by
induction on the number of elements in a fixed generating set for $M$, which we
informally describe: By definition, $x \in I M$ can  be written, for some
positive integer $k$ and elements $r_i \in R$ and $w_i \in M$,
$1 \leq i \leq k$, as $x = r_1 \cdot w_1 + r_2 \cdot w_2 + \cdots +
r_k \cdot w_k$. What's more, by definition of a finite generating set $u_{1},
\dots, u_{n}$ for $M$, each $w_i$ equals $c_{i1} \cdot u_1 + c_{i2} \cdot u_2 +
\cdots + c_{in} \cdot u_n$ for appropriate choices of $c_{ij} \in R$. Hence,
inductively applying associativity, 
\begin{align*}
  x = \sum_{i=1}^{k} r_i \cdot w_i
    = \sum_{i=1}^{k} r_i \cdot \left(\sum_{j=1}^n c_{ij} \cdot  u_j\right)
    = \sum_{j=1}^{n}\sum_{i=1}^{k} r_i \cdot (c_{ij} \cdot u_j) 
    = \sum_{j=1}^{n}\left(\sum_{i=1}^{k} c_{ij} \cdot r_i\right) \cdot u_j.   
\end{align*}
By the absorption property of ideals, each  $c_{ij}  \cdot r_i$ is in $I$, and since ideals
are closed under addition, we inductively conclude that the coefficient $a_j
:= c_{1j} \cdot  r_1 +  c_{2j} \cdot  r_2 +  \cdots c_{kj}\cdot r_k$ of $u_j$
is also in $I$.

\subsection{The Formal Proof}
We now outline our formal proof of Nakayama's lemma, and  proceed
by way of induction on the number of elements in a fixed generating set for the finitely generated $R$-module
$M$. The base case in this situation is when \(M\) requires no generators, so $M$
consists solely of the empty scalar combination, i.e., the empty sum, which, by
convention, is the zero element. In other words, $M = 0$ by assumption, and
this is also precisely the conclusion of Nakayama's Lemma.  Hence the statement holds
trivially, without using the hypothesis that $M = \mathfrak{m} M$.

@@

Now, fix an $R$-module $M$ generated by $u_1, \ldots, u_{n+1} \in M$. By
assumption, \(M = \mathfrak{m} M\), and in particular, the generator $u_1$ is
an element of the submodule $\mathfrak{m} M$. The lemma described in \Cref{formal-statement: SS}
guarantees the existence of ring elements $a_1, \ldots, a_n \in \mathfrak{m}$
for which 
\[u_{1} = a_{1} \cdot u_{1} + a_2 \cdot u_2 + \cdots + a_{n+1} \cdot u_{n+1}.\]
Collecting the \(u_{1}\) terms on the left-hand side of the equation, calling on 
the existence of additive inverses in $R$, and distributivity of scalar
multiplication for modules, we see that
\begin{equation}\label{equation1: e}
(1 - a_{1}) \cdot u_{1} =
  a_{2} \cdot u_{2} + \cdots + a_{n+1} \cdot u_{n+1}\text{.}
\end{equation}

In Coq, we work backward to establish \eqref{equation1: e}. The comment following
each tactic appearing in the code below offers a brief indication of its effects on the goal.
\begin{minted}[linenos,firstnumber=157,stepnumber=5]{coq}
  assert ((Rone [+] Rminus a1) <.> u1 =M= (* (1 - a1) u1 = ... *)
    linear_combin Madd Mzero action coeffs' generatingSet').
  { setoid_rewrite (module_distrib_Radd Radd Rmul Rone
      Mequiv Madd Mzero Mminus action). (* 1 u1 + (-a1) u1 = ... *)
    setoid_rewrite (module_Rone Radd Rmul Rone
      Mequiv Madd Mzero Mminus action). (* u1 + (-a1) u1 = ... *)
    setoid_rewrite (module_minus_l Requiv Radd Rzero Rminus Rmul Rone
      Mequiv Madd Mzero Mminus action). (* u1 - a1 u1 = ... *)
    symmetry. (* ... = u1 - a1 u1 *)
    apply (group_move_r Mequiv Madd Mzero Mminus). (* ... + a1 u1 = u1 *)
    symmetry. (* u1 = ... + a1 u1 *)
    assumption. }
\end{minted}

Since a proper ideal contains no units and $a_1$ is taken to be in
$\mathfrak{m}$, it is not a unit. Hence, by the formalized statement mentioned
early in \Cref{formal-proof: S}, $1 - a_{1}$ is a unit of $R$. Let $b_{1} \in R$ denote its 
multiplicative inverse, so that $b_1 \cdot  (1 - a_{1}) = 1$. Multiplying this
element on the either side of \eqref{equation1: e}, we deduce the following.  
\begin{align}
\label{simplify-line-1}
b_1 \cdot \left((1 - a_{1}) \cdot u_{1} \right)
  & = b_1 \cdot (a_{2} \cdot u_{2} + \cdots + a_{n+1} \cdot u_{n+1}) \\ 
\label{simplify-line-2}
\left( b_1 \cdot (1 - a_{1})\right) \cdot u_{1}
  & = b_1 \cdot (a_{2} \cdot u_{2}) + \cdots + b_1 \cdot (a_{n+1} \cdot u_{n+1}) \\ 
   \label{simplify-line-3}
u_1 = 1 \cdot u_{1}
  & = (b_1 \cdot  a_{2}) \cdot u_{2} + \cdots + (b_1  \cdot  a_{n+1}) \cdot u_{n+1} 
\end{align}
\begin{minted}[linenos,firstnumber=181,stepnumber=5]{coq}
  pose proof
    (local_comm_ring_sub_1_nonunit Requiv Radd Rzero Rminus Rmul Rone
      R_local x1 Hx1_nonunit) as H1ma1_unit.
  (* Getting to u1 = (b1 coeffs') . generatingSet *)
  inversion_clear H1ma1_unit as [b1 Hb1]. (* 1 - a1 is a unit *)
  setoid_rewrite (commutative Requiv Rmul) in Hb1. (* b1 (1 - a1) = 1 *)
\end{minted}

This block of code has the effect of naming \(b_{1}\) as the inverse of the
unit \(1 - a_{1}\), and storing in the hypothesis \texttt{Hb1} the equation
\(b_{1} \cdot (1 - a_{1}) = 1\). The remainder of the code goes on to multiply on the
left by \(b_{1}\) in the equation \texttt{H} in order to establish  \eqref{simplify-line-1}.
Going on and applying the axioms of a module over a commutative ring, along with hypothesis \texttt{Hb1}, we can
simplify the left hand side of \texttt{H} to be just \(u_{1}\), as in \eqref{simplify-line-3}.
\begin{minted}[linenos,firstnumber=last,stepnumber=5]{coq}
  apply (module_op_l Requiv Mequiv action) with (r:=b1) in H.
  setoid_rewrite <- (module_distrib_Rmul Radd Rmul Rone
    Mequiv Madd Mzero Mminus action) in H. (* H: (b1 (1 - a1)) u1 = ... *)
  setoid_rewrite Hb1 in H. (* H: (1) u1 = ... *)
  setoid_rewrite (module_Rone Radd Rmul Rone
    Mequiv Madd Mzero Mminus action) in H. (* H: u1 = ... *)
\end{minted}
In particular, the generator $u_1$ can be written as a scalar combination of
the $n$ generators $u_2, \ldots, u_{n+1}$, and is superfluous. Hence $M$ can be
generated by $n$ elements, and $M=0$ by the inductive hypothesis. 
Though showing that the inductive hypothesis holds is more technical in Coq than in this
informal proof, the extra steps are essentially bookkeeping.

\section{Conclusion}
Our work here is evidence that Coq is amendable to the 
formalization of abstract algebraic notions from commutative algebra, which are often non-discrete, and their theory. 
There are several natural directions in which to move from here, a first  step being to further develop the formal theory of quotient rings, and to formalize the isomorphism theorems for rings.
